---
title: "Week 01: Lab"
output: html_notebook
---

# Lab: Squared Away

## Step 0ne
- Installation/Setup of class tools:
    - Log into our VMs
    - R, 
    - RStudio, 

```{r message=FALSE, include=FALSE}

install.packages("openxlsx")
install.packages("tidyverse")

```

Now that we have installed some things, let's see what we have.

```{r}
data.frame(installed.packages())
```

## Step Two
- RStudio Intro
    - What is an IDE?
    - RStudio Tour
    - Using Help
    - **?? Questions ??**
- Things you can do with R
    - A rather useful calculator
    - View data
    - Graph data
    - Rule you data!

### R: A rather useful calculator

```{r}

## How to use R as a calculator.


```

### View data

```{r}

## What data would you like to view?
## What kinds of data are in it?
data()

```


### Graph Data

**Question:** Does the sepal length of an iris correlate with the petal length?
**Answer:** Yes, with an interesting step function in the middle.

```{r}
plot(iris$Sepal.Length, iris$Petal.Length)
```


### Rule Data

```{r}
plot(iris)
```

## Step Three
- Questions?

## Step Four
- Compare to Excel