---
title: "Week 01: Lab"
output: html_notebook
---

# Lab: Squared Away

## Step Zero
- Write some code!
- "Hello World" is a time-honored first step you should take in ANY programming
  language.
  
```{r}
## Hello World!


```


## Step 0ne
- Installation/Setup of class tools:
    - Log into our VMs
    - R, 
    - RStudio, 

```{r message=FALSE, include=FALSE}

install.packages("openxlsx")
install.packages("tidyverse")

```

Now that we have installed some things, let's see what we have.

```{r}
data.frame(installed.packages())
```

## Step Two
- RStudio Intro
    - What is an IDE?
    - RStudio Tour
    - Using Help
    - **?? Questions ??**
- Things you can do with R
    - A rather useful calculator
    - View data
    - Graph data
    - Rule you data!

### R: A rather useful calculator

```{r}

## How to use R as a calculator.


```

### View data

```{r}

## Which of these? interests you?
data()

```


### Graph Data

Let's use the "iris" data to look at how easy it is to use R.

```{r}
data(iris)
head(iris, 15)
```


**Question:** Does the sepal length of an iris correlate with the petal length?
**Answer:** Yes, with an interesting step function in the middle.

```{r}

## We can use the plot function to explore relationships in our data.


```


### Rule Data

```{r}

## We can also just go bananas with plot and plot everything against everything.
## Plot ALL THE THINGS!

```

## Step Three
- Questions?
- Concerns?

```{r}
## We can all share code via Slack!


```


## Step Four
- Compare to Excel

## Step Five
- Congratulations
- I don't want to overwhelm you. Today, we are just getting our feet wet.
- How to submit this to me.
- Questions?