---
title: "Week 02: Vectors"
output:
  html_notebook:
    toc: true
    toc_depth: 2
---

# Introduction
- Welcome back, blah, blah, blah.
- If you haven't already done so, please put your artistically designed name-banner-thing out where I can see it to help me call you by your actual name.
- Tonight's goals: 
    - Review what we learned last week
    - Introduce a core R concept called the vector
    - Getting Help
    - Discuss some sampling concepts
     - But it all starts with a little detour down data type lane

# Last Week Tonight (Without John Oliver)

## Using R As A Calculator

```{r}
## Example: Create two variables, a and b.
## Then add them together.
a <- 1
b =  2
c <- a + b
c
```

- Although `<-` is the canonical assignment operator, `=` works.
- R, and most other programming languages _are_ case sensitive.
    - Some languges such as SAS and SQL are not.
    - R, Python, C, etc are case sensitive.
```{r eval=FALSE}
# Do you understand why this fails?
A
```
    
- If you see code using `<<-`, and don't know what terms like global variable and scoping mean - run away.
- *Note:* I am giving you my lecture notes, to help you learn both R and Markdown.
    - Render the output, hit the Preview button above.
    - Are there any questions about how to run a code chunk?
    - Are there any questions about the RStudio IDE?
    - Does anyone want to order some pizza?

# Data Types (Which We Want To Discuss Today)

- Numeric
- Character
    - Factor
- Logical

## Numeric

```{r eval=FALSE}
## Create a variable called my_age. 
## Assign it a value of your age (or how old you feel).
my_age <-

```

Arithmetic is straightforward
- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`


## Character/Factor

- Mostly, things that are not numbers.
- Names, Gender, Race, Favorite Color, Address of nearest McDonald's

```{r eval=FALSE}
## Create a variable called my_name.
## Assign it a value of your name.

my_name <-
    
```

- A factor is a special type of character, which we will learn more about, which
  has repetitive values.
- Sometimes referred to as a string variable. I tend to use these terms
  interchangeably.
- Example: Each of you has, I presome, a different home address. That
  information would be stored as a character.
- But things like gender, race, and years in school would be stored as factors.
    - We will do more with this, I promise.
    - If we don't, blame someone else.

```{r}
## Addition doesn't work with characters. But you can paste.
## This is called concatenation in many other languages.
foo <- "a"
bar <- "b"

paste(foo, bar, sep = " | ")

```

## Logicals
- Logical | Boolean | Whatever
- These are "protected" words - if you spell them right.
- True must be spelled `TRUE` or `T`.
- False must be spelled `FALSE` or `F`.
- I prefer to spell them out.
    - Even if you are a terrible typist, the additional effort is unlikely to 
      harm you, much.
- All other spellings are just chacters.
    - `True`, `False` are not `TRUE`, `FALSE`.
- You can apply Boolean Logic in your code.
- The following code chunks use the variables `a`, `b`, `c`, `foo`, and `bar` 
  defined above. If you haven't run those code chunks, do that before running 
  the code below.

```{r}
## Tests of equality
a == b
a != b
```

```{r}
## More tests of equality
foo == bar
foo != bar
```

```{r}
## We can do other tests.
a > b
c <= a
```

- Equals: `==`
    - (Remember, =, is a just a hipster assignment operator.)
- Less Than: `<`, `<=`
- Greater Than: `>`, `>=`
- Not Equal: `!=`
- All logic tests return the value of TRUE, FALSE, or NA.
- Useful things to know!
    - TRUE has a numeric value of 1.
    - FALSE has a numeric value of 0.

```{r}
a <- 1
b <- 2
a > b
```

Some functions return logicals.

```{r}
is.numeric(a)
is.character(foo)
is.numeric(bar)
```

# Vectors
- At long last - the point of this lecture!
- Everything in R is an "object". One type of "object" is called a vector.
- In fact, we have already created many vectors, I just called them variables.
    - `a`, `b`, `c`, `foo`, and `bar` are all vectors, length == 1.

```{r}
## Play around with this. Using the other variables we have made, do you ever
## get different results?
is.vector(foo)
length(a)
```

Fortunately, vectors can have lengths longer than one.

```{r}
## We can make a vector of numbers as a sequence.
a <- 1:15
a

## Feel free to play the song.
b <- c(8,6,7,5,3,0,9)
b

## Question: How can we find out what "c" means and how to use it?
## I'm looking for two answers. Via the IDE and via the REPL.

## Or we can create a vector as output from another function:
c <- sample(1:6, 33, replace = TRUE)
c
```

Please note that these vectors are shown as part of or global environment. From
there, we can learn more about them.

```{r}
## Can you tell me the length of b and c?
length(a)
```

## OMG Excel
- We can do something similar in Excel. Let's open Excel and create these vectors
  in Excel.
  
## Atomic
- Welcome back. R kept the coffee warm and tasty.
- In R - vectors are sometimes referred to as "atomic vectors". In R a vector is 
  defined as a one dimensional (length) object of the same type. Above - we 
  created numeric vectors. But a vector can be made of any SINGLE data type.

```{r}
## This is both silly and rather pointless, but it works.
foo <- c("a", "b", "c", "three")
length(foo)
```

```{r}
## This also works works, but it does something.
## Any idea what this does?
bar <- c("a", "b", 8)
bar
```

Fortunately, there is a lot we can do with a vector.
- Math
- Functions
- Plotting

```{r}
## Math
## Compare a to a+1
a
a + 1
```

```{r}
## Functions
## Many functions in R can accept a vector as a parameter.
## Remember, your results may differ a little because we built this randomly.
table(c)
```

```{r}
## Plotting
## Again - your results may differ.
plot(as.factor(c))
```

*Question:* What kind of a plot is this and what kind of data would we use it 
on?

- Oops. This might be a good time to stop and discuss what a factor is.
- These are a special type of character vector.
- They have a finite number of types - like grade level, favorite football team,
  etc. They are not endlessly variable (address) or continuous (height).
- The `plot` function operates based on the TYPE of data passed to it. Because 
  `c` is a vector of numbers - it would try to scatter plot them. But, in this 
  case, `c` is better represented by a histogram and the way to trigger that is 
  to tell `plot` to convert c to a factor.

## Some Functions

Time for a moment of independent study.

| Mathematical              | Useful                |
|:--------------------------|:----------------------|
| Average: `mean()`         | # Items: `length()`   |
| Standard Deviation: `sd()`| Crosstab: `table()`   |
| Variance: `var()`         | What?: `summary()     |
| Smallest: `min()`         | Crosstab: `plot()`    |
| Biggest: `max()`          | Histogram: `hist()`   |
|                           | Barplot: `barplot()`

*Questions?* Speak Now!

## Basic Indexing on Vectors
- Sometimes we don't want to use _every_ entry in a vector. Sometimes, we only
  want to use a subset of them.

```{r}
## Indexing!
## Always using the square bracket operator
c[1:3]
```

- This is nice for demonstrative purposes.
- Much more useful is to apply boolean logic to vectors. 

## Boolean Logic On Vectors
We can even do boolean tests on our vectors:

```{r}
c > 3
```

```{r}
table(c > 3)
```

We can also apply boolean tests to get only those values we want!

```{r}
c[c > 3]
```

We can use Boolean logic with other data types, like characters.

```{r}
gender <- sample(c("Foo", "Bar", "Baz"), 24, replace = TRUE) 
gender[gender != "Baz"]
```

And , it might not seem obvious, but this works too!

```{r}
## Oh yes, we can!
gender[gender > "Baz"]
```

## Getting Help
- ?foo
- Code Completion
- RStudio help window
- [Stack Overflow](http://stackoverflow.com/questions/tagged/r)
- Google: Pre-pend search query with "R"
- [Quick-R (Stat Methods)](https://www.statmethods.net/index.html)
- [RStudio Cheatsheets](https://www.rstudio.com/resources/cheatsheets/)

## Difference Between Sample and Population

The vector, `c`, is a random sample of values between 1 and 6. 

```{r}
## Let's create a new sample
c <- sample(1:6, 20, replace = TRUE)
```

Each entry in this vector has an equally random chance of being 1, 2, 3, 4, 5, 
or 6. In theory, the odds for each number are equal and are equal to:

```{r}
1/6*100
```

*Question:* So what is the difference between a sample and a population?