---
title: "Transformations & EDA"
author: "Andy Choens"
date: "Week 03"
output:
  ioslides_presentation:
    widescreen: TRUE
    smaller: TRUE
#runtime: shiny
---


## Load Your Libs!

```{r}
library(tidyverse)
library(readxl)
```


## Project

- Create a folder to hold all of your code/projects for this course.
- I would put your lecture notes into a folder each week (for
downloaded data, scripts, etc.)
- I call my folder `Intro To Data`.
- You can call your folder anything you please.
- In this folder, you should have a folder for each lecture and each lab.
- I recommend something obvious like Lecture 03. But again, you do you.


## MPG

- We have used a dataset called mpg for a couple of porpoises:

```{r echo=FALSE}
data(mpg)
head(mpg) %>% knitr::kable()
```

- This dataset has `r nrow(mpg)` rows of data and `r ncol(mpg)` columns.
- These "toy" datasets are cute, but let's go get a real data set.

## MPG, But Bigger (1)

### [https://www.fueleconomy.gov/](https://www.fueleconomy.gov/)

The US Federal Government still publishes annual estimates for the mileage of 
cars sold in the US each year. Let's download the 2020 numbers.

- At the bottom of the page: "Download EPA's MPG Ratings"
- We found the data. What else should we be looking for?


## MPG, But Bigger (2)

Always look for documentation to help you understand the: 

- content of the data (columns)
- sampling/inclusion/exclusion criteria (as relevant/appropriate)
- how much data is there (this goes back to 1978)
- caveats, changes in the data over time.
- Any aggregations we can use to validate our analyses.

We will download the 2020 data. But down't go clickig anything just yet.

## MPG, But Bigger (3)

You are using a project now, aren't you?

```{r eval=FALSE}
## Use variables. It is easier to read and more flexible.
data_url <- "https://www.fueleconomy.gov/feg/epadata/20data.zip"
local_file <- "mpg20.zip"

download.file(url = data_url, destfile = local_file)
unzip(local_file)
dir(pattern="xlsx")
```

- Feel free to open the data in Excel
- What do you see?
- What would you like to know?

<span style="color:red">STOP ME NOW if you are having trouble!</span>

## MPG, But Bigger (3)

Let's import the data and look at it.

```{r}
data_file <- dir(pattern="xlsx")
Mpg <- read_excel(data_file)
head(Mpg)
```

## STOP!

## Let's talk about grain.

- Wheat
- Quinoa
- Spelt
- Oat

No, seriously, is this data "tidy" and what is the grain of this data?


## Grain, Why It Matters

Could we use this data to discuss the average mileage:

- of all cars on the road?
- of all cars sold in 2020

## Column Names (1)

- Those column names are not going to be easy to work with.
- But they do contain useful info.
- It is a bad practice to throw out information.
- So, we will create an attribute of Mpg to hold the originals.

```{r}
names(Mpg)

attributes(Mpg)$labels <- names(Mpg)
```


## Column Names (2)

- Now we can fix those column names.
- There are 162 columns in this data. We _could_ change the columns this way:

```{r eval=FALSE}
names(Mpg) <- c("ModelYear", "MfrName", "Division", "Carline")
```

OMG, you will be here all night doing this. Let's automate it.

## Column Names (3)

```{r}
col_names <- names(Mpg)

## Generic Changes
col_names <- gsub("- Conventional Fuel", "", col_names, fixed=TRUE)
col_names <- gsub("- Alternative Fuel", "", col_names, fixed=TRUE)
col_names <- gsub(" ", "", col_names, fixed=TRUE)
col_names <- gsub(".", "", col_names, fixed=TRUE)
col_names <- gsub("?", "", col_names, fixed=TRUE)
col_names <- gsub(",", "", col_names, fixed=TRUE)
col_names <- gsub("-", "", col_names, fixed=TRUE)
col_names <- gsub("%", "", col_names, fixed=TRUE)
col_names <- gsub("#", "N", col_names, fixed=TRUE)

# Specific Tweaks
col_names[6] <- "Indexed"
col_names[10] <- "CityFECF"
col_names[11] <- "CityFECF"
col_names[12] <- "CombFECF"
col_names[13] <- "CityUnadjFECF"
col_names[14] <- "HwyUnadjFECF"
col_names[15] <- "CombUnadjFECF"
col_names[16] <- "CityUnrdUnadjFECF"
col_names[17] <- "HwyUnrdUnadjFECF"
col_names[18] <- "CombUnrdUnadjFECF"
col_names[19] <- "Guzzler"
col_names[30] <- "MaxEthanolGasoline"
col_names[31] <- "MaxBiodiesel"
col_names[32] <- "Range1CF"
col_names[33] <- "FuelUsageCF"
col_names[151] <- "MoneySaved5Year"
col_names[152] <- "MoneySpend5Year"
col_names[157] <- "ChargeTimeat240volts"
col_names[158] <- "ChargeTimeat120volts"
col_names[159] <- "PHEVRange"

## How can we change:
## [46] "EPA Calculated Annual Fuel Cost  -----  Annual fuel cost error. Please revise Verify."  

names(Mpg) <- col_names
```

## EDA

- We are going to do a choose your own adventure here.
- The goal: 
   - Look for patterns. 
   - Look for centrality. 
   - Look for outliers.
- In doing this, we will introduce some graphs AND some transformations.
- This will serve as your intro to dplyr and a review of ggplot.
- You will never stop learning about programming.
   - Never